Deep Learning Project

### 2024-10-11 project/paper-final.tex
Added final latex and PDF's.

\textcolor{red}{
\section {Motivation}
The primary motivation for attempting to reproduce the experiment from the Szegedy paper was to apply TensorFlow to a practical problem, such as image classification, using a realistic dataset. The detailed documentation of the network in the Szegedy paper greatly facilitated its TensorFlow implementation. Recognizing that the operational environment would be constrained (i.e., running the reproduction on a desktop PC), we also saw this as an opportunity to evaluate the network's performance on a smaller version of the ImageNet dataset. This project provided a hands-on experience with the challenges of training a neural network. Additionally, it offered valuable experience in writing TensorFlow code for a complex neural network with branching and joining structures, a significant leap from our prior work with simpler sequential networks. For comparison, we also experimented with a sequential network to highlight the differences in complexity and performance.}

\textcolor{red}{
\section{Inspiration}
Two code bases served as the inspiration for our Tensorflow implementation of Inception. The image pre-processing, particularly regularization techniques were guided by \cite{dishasai_multiclass_classification}. The Inception layer implementation was inspired by \cite{googlenet_inception_module}. Lastly, the graphs in Figures \ref{fig:googlenet-simple-accuracy} and \ref{fig:googlenet-complex-accuracy} were inspired by \cite{datacamp_cnn_tensorflow}.}

@@

\textcolor{red}{
\subsection{Compute Resources}
The networks created for this project were trained and tested using GPU and CPU resources for comparison purposes. The GPU is an NVIDIA Quadro P2000 GPU having 5GB of dedicated memory. The CPU is an Intel Core i9-10850K 3.6GHz CPU with 20 logical cores. }

@@

\subsection{Replicating the GoogLeNet Network}
The GoogLeNet architecture is explained in Table 1 of the Szegedy paper is reproduced in Figure \ref{fig:googlenet-arch}. \textcolor{red}{The network is 22 layers deep, incorporating branching and parallel convolutions via Inception modules.} The introduction of Inception layers having parallel convolutional filters is the primary difference from typical sequential networks. The replication attempt, when trained for 30 epochs, achieved a maximum accuracy of 60.3\%. This is quite improved over the simple model. The other impressive aspect is that it only took 14 epochs to arrive at the maximum accuracy as shown in Figure \ref{fig:googlenet-complex-accuracy}. The conclusion is that the architecture does provide marked improvements over the simpler, more naive model described in \ref{subsec:simple_model}.

@@

\textcolor{red}{\subsubsection{Computational Demand}}
\textcolor{red}{One of the challenges we encountered was the high computational demand of training, even on the subset of ImageNet images we used.  The training performance is summarized in Table \ref{tab:training_times}. It is interesting that the training times for the two models were nearly identical when using a \textit{single GPU} but differed by a factor of 1.7 when using \textit{20 logical CPU's.} This drove home one of the reasons GPU's have become so popular. From the training times, it should be readily apparent why a subset of the ImageNet images were chosen and that training the whole dataset would take an inordinate amount of time with the compute resources we had access to. To further reduce the time, the learning rate was reduced to 0.0001, and batch size was reduced to 16. Prefetching was also implemented to improve performance\cite{tensorflow_data_performance}.} \textcolor{red}{\sout{Although GoogLeNet is more efficient than many of its contemporaries, training still required GPU resources. We used data augmentation techniques, such as random cropping, random horizontal flipping, and random zoom to avoid overfitting\cite{dishasai_multiclass_classification}. We also needed to reduce the learning rate to 0.0001, reduced batch size to 16, and enabled memory growth\cite{tensorflow_memory_growth} so as not to overwhelm the GPU. Prefetching was also implemented to improve performance\cite{tensorflow_data_performance}. The training performance is summarized in Table \ref{tab:training_times}. From the training times, it should be readily apparent why a subset of the ImageNet images were chosen and that training the whole dataset would take an inordinate amount of time with the compute resources we had access to.}}

@@

\textcolor{red}{\subsubsection{GPU Setup}}
\textcolor{red}{The setup required to run Tensorflow code on a GPU on Windows is not trivial. Fortunately two good resources provided instructions on accomplishing this task \cite{abulhawa_tensorflow_gpu_wsl2_2023} \cite{towardsdatascience_jupyter_wsl2_2023}.}

\textcolor{red}{\subsubsection{Overfitting}}
\textcolor{red}{Overfitting was also an issue. We used data augmentation techniques, such as random cropping, random horizontal flipping, and random zoom to reduce overfitting\cite{dishasai_multiclass_classification}.}

@@

\textcolor{red}{
\subsection{Influence on Present Technology}
Image recognition technology has become ubiquitous and is now readily accessible to the masses. For example, Google Photos can organize images based on their content \cite{google_photos_updates_2023}, and one can take a picture of a food label written in a foreign language and translate it using Google Translate \cite{google_translate_support_2024}. The success of neural networks can be largely attributed to the influence of Szegedy's work and its experiments. However, the original Szegedy experiments required substantial computing power, making them impractical for resource-constrained environments. Since 2015, there have been significant advancements in creating deep networks that can efficiently run on memory-limited devices, such as smartphones and automobiles. MobileNet, as described in \cite{howard2017mobilenets}, is a class of efficient models designed specifically for mobile applications. It builds upon key innovations from Inception, including multi-scale feature extraction, the use of 1x1 convolutions for dimensionality reduction, and the stacking of layers.}

>>>>

\section {Motivation}
The primary motivation for attempting to reproduce the experiment from the Szegedy paper was to apply TensorFlow to a practical problem, such as image classification, using a realistic dataset. The detailed documentation of the network in the Szegedy paper greatly facilitated its TensorFlow implementation. Recognizing that the operational environment would be constrained (i.e., running the reproduction on a desktop PC), we also saw this as an opportunity to evaluate the network's performance on a smaller version of the ImageNet dataset. This project provided a hands-on experience with the challenges of training a neural network. Additionally, it offered valuable experience in writing TensorFlow code for a complex neural network with branching and joining structures, a significant leap from our prior work with simpler sequential networks. For comparison, we also experimented with a sequential network to highlight the differences in complexity and performance.

\section{Inspiration}
Two code bases served as the inspiration for our Tensorflow implementation of Inception. The image pre-processing, particularly regularization techniques were guided by \cite{dishasai_multiclass_classification}. The Inception layer implementation was inspired by \cite{googlenet_inception_module}. Lastly, the graphs in Figures \ref{fig:googlenet-simple-accuracy} and \ref{fig:googlenet-complex-accuracy} were inspired by \cite{datacamp_cnn_tensorflow}.

@@

\subsection{Compute Resources}
The networks created for this project were trained and tested using GPU and CPU resources for comparison purposes. The GPU is an NVIDIA Quadro P2000 GPU having 5GB of dedicated memory. The CPU is an Intel Core i9-10850K 3.6GHz CPU with 20 logical cores.

@@

\subsection{Replicating the GoogLeNet Network}
The GoogLeNet architecture is explained in Table 1 of the Szegedy paper is reproduced in Figure \ref{fig:googlenet-arch}. The network is 22 layers deep, incorporating branching and parallel convolutions via Inception modules. The introduction of Inception layers having parallel convolutional filters is the primary difference from typical sequential networks. The replication attempt, when trained for 30 epochs, achieved a maximum accuracy of 60.3\%. This is quite improved over the simple model. The other impressive aspect is that it only took 14 epochs to arrive at the maximum accuracy as shown in Figure \ref{fig:googlenet-complex-accuracy}. The conclusion is that the architecture does provide marked improvements over the simpler, more naive model described in \ref{subsec:simple_model}.

@@

\subsubsection{Computational Demand}
One of the challenges we encountered was the high computational demand of training, even on the subset of ImageNet images we used.  The training performance is summarized in Table \ref{tab:training_times}. It is interesting that the training times for the two models were nearly identical when using a \textit{single GPU} but differed by a factor of 1.7 when using \textit{20 logical CPU's.} This drove home one of the reasons GPU's have become so popular. From the training times, it should be readily apparent why a subset of the ImageNet images were chosen and that training the whole dataset would take an inordinate amount of time with the compute resources we had access to. To further reduce the time, the learning rate was reduced to 0.0001, and batch size was reduced to 16. Prefetching was also implemented to improve performance\cite{tensorflow_data_performance}.

@@

{\subsubsection{GPU Setup}
The setup required to run Tensorflow code on a GPU on Windows is not trivial. Fortunately two good resources provided instructions on accomplishing this task \cite{abulhawa_tensorflow_gpu_wsl2_2023} \cite{towardsdatascience_jupyter_wsl2_2023}.

\subsubsection{Overfitting}
Overfitting was also an issue. We used data augmentation techniques, such as random cropping, random horizontal flipping, and random zoom to reduce overfitting\cite{dishasai_multiclass_classification}.

@@

\subsection{Influence on Present Technology}
Image recognition technology has become ubiquitous and is now readily accessible to the masses. For example, Google Photos can organize images based on their content \cite{google_photos_updates_2023}, and one can take a picture of a food label written in a foreign language and translate it using Google Translate \cite{google_translate_support_2024}. The success of neural networks can be largely attributed to the influence of Szegedy's work and its experiments. However, the original Szegedy experiments required substantial computing power, making them impractical for resource-constrained environments. Since 2015, there have been significant advancements in creating deep networks that can efficiently run on memory-limited devices, such as smartphones and automobiles. MobileNet, as described in \cite{howard2017mobilenets}, is a class of efficient models designed specifically for mobile applications. It builds upon key innovations from Inception, including multi-scale feature extraction, the use of 1x1 convolutions for dimensionality reduction, and the stacking of layers.
